---
title: "Tidyverse Fun - Part 1"
description: |
  First part in a series of doing useful tasks with 
  the `Tidyverse`
author:
  - name: Shamindra Shrotriya
    url: https://www.shamindras.com/
date: 07-15-2019
preview: images/logo.png
slug: shrotriya2019tidyfunpt1
bibliography: ../../refs.bib
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup_01_01, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prev_img, preview=TRUE, results='hide'}
IMAGES_PATH <- here::here("_posts",
                          "2019-07-25-shrotriya2019tidyfunpt1",
                          "images")
knitr::include_graphics(path = fs::path_join(parts = c(IMAGES_PATH, 
                                                       "tidyverse.png")))
```

## Task 1: Generating Oxford Comma Triples

### The central problem

Based on a fun conversation with my statistics cohort 
over dinner we got to discussing the famous *Oxford Comma* [@wiki_serial_comma] (or *Serial Comma* depending on your
persuasion). I've never really adopted the use but my
friends made a compelling argument on it's general lack
of ambiguity in phrases compared to not using it.

We will use the Oxford comma on the famous *ambiguous*
phrase (here used without the Oxford Comma 
before 'shoots'):

> Eats, shoots and leaves

After adding in the Oxford Comma this would become:

> Eats, shoots, and leaves

**Goal:** A fun experiment would be to generate *all permutations* of this phrase with and without 
the Oxford Comma.

### Generating all word-triple permutations the `tidy` way

Let's load our required packages

```{r setup_oxc_01_01, echo=TRUE, cache=TRUE}
library(magrittr)
library(tidyverse)
library(glue)
```

```{r setup_oxc_01_02, echo=TRUE, cache=TRUE}
# Define our unique GLOBAL word values
WORD_VALS <- c("eats", "shoots", "leaves")
```

Generate all unique 3-word permutations without replacement from the three unique words. We'll create a helper
function to check that a vector of words is unique.

```{r setup_oxc_01_03, echo=TRUE, cache=TRUE}
is_unq_perm <- function(word1, word2, word3){
    words_vec <- c(word1, word2, word3)
    base::return(length(words_vec) - length(unique(words_vec)) == 0)
}
```

Here we simply generated every possible 3 word triple
with replacement using the `tidyr::crossing` function.
We simply filter these $3^3 = 27$ triples for unique
triples using the `is_unq_perm` function applied
row-by-row using `purrr::pmap_lgl`. The `_lgl` simply
returns a `TRUE/FALSE` logical value as intended by
the applied function.

<aside>
This generates a *Cartesian product* of all the 3 word
triples
</aside>

```{r setup_oxc_01_04, echo=TRUE, cache=TRUE}
# Generate the unique word-triples
all_perms <- tidyr::crossing(word1 = WORD_VALS, 
                             word2 = WORD_VALS, 
                             word3 = WORD_VALS) %>% 
                dplyr::mutate(.data = ., 
                              is_unq_perm = purrr::pmap_lgl(.l = ., 
                                                            is_unq_perm)) %>% 
                dplyr::filter(.data = ., is_unq_perm) %>% 
                dplyr::select(-is_unq_perm)

# Display output in a nice centered table
all_perms %>% 
  knitr::kable(x = ., align = 'c',
               col.names = c("Word 1", 
                             "Word 2", 
                             "Word 3"))
```

Great - that part is done! Now we just need to generate
for each triple of words an oxford comma and non-oxford
comma version. This is done easily using the amazing
`glue` package as seen below:

```{r setup_oxc_01_05, cache=TRUE, echo=TRUE}
exprs <- all_perms %>%
          dplyr::mutate(non_oxford_comma = 
                          glue::glue_data(.x = ., 
                          "{word1}, {word2} and {word3}"), 
                        oxford_comma = 
                          glue::glue_data(.x = ., 
                          "{word1}, {word2}, and {word3}")) %>% 
          dplyr::select(non_oxford_comma, oxford_comma)
```

We can display the side-by-side output of the Non-Oxford
Comma vs. Oxford comma for the $6$ generated triples as
follows:

```{r setup_oxc_01_06, cache=TRUE, echo=TRUE}
# Display output in a nice centered table
exprs %>% 
  knitr::kable(x = ., 
               align = 'c', 
               col.names = c("Non-Oxford Comma", 
                             "Oxford Comma"))
```

So there you have it. Have fun generating your
own version of Oxford Comma triples to engage
in civil discussion with your fellow grammar 
focused friends `r emo::ji("smile")`.

## Task 2: Generating numbered BibTeX entries

### The central problem

In this case I needed to generate several BibTeX entries of the form:

```markup
@misc{doe2019_lec1,
author        = {Doe, John},
title         = {Lecture Note 1 - STAT10A},
month         = {March},
year          = {2018},
url           = {https://statschool/~doe/stats10A/Lectures/Lecture01.pdf},
}
```

Specifically 30 such sequential entries to create for lectures `1-30`. Rather
than do this manually, I realized that
this would be fun scripting exercise with using the
`tidyverse` packages `glue`, `purrr`, and `stringr`.

**Goal:** Create 30 such BibTeX entries and print to 
the console to directly-copy paste to my BibTeX file.

### The `tidy` approach

First step is to write a function that takes a lecture number (integer)
as an input and then outputs a single BibTeX entry for that lecture.

```{r bbtex_01_01, eval=TRUE, echo=TRUE, cache=TRUE}
# Generate BibTeX entry for a single lecture number
get_lec_bibtex <- function(lec_num){
  # Get the 2 character padded lecture number i.e. 1 -> "01"
  lec_num_pad <- stringr::str_pad(string = lec_num, width = 2, 
                                  side = "left", pad = "0")
  
  # Construct the BibTeX entry
  out_bbtex_str <- glue::glue(
    "@misc{doe2019_lec<lec_num>,
    author = {Doe, John},
    title  = {Lecture Note <lec_num> - STAT10A},
    month  = {March},
    year   = {2018},
    url    = {https://www.hpg/~doe/st10A/lecs/lec<lec_num_pad>.pdf}}",
    .open = "<", 
    .close = ">")

  base::return(out_bbtex_str)
}
```

Note that by default `glue` allows you to substitute input text in
between `{` and `}` markers. However BibTeX entries *already
have* literal default `{}` tags that we need to include in
our function output. Rather than escaping them the `glue` 
package conveniently allows us to change the default opening 
and closing markers `r emo::ji("hundred")`! We simply set 
these to be angle brackets `< >` using the `.open` and `.close` 
options above.

<aside>
Luckily we don't have *literal* angle brackets in our BibTeX
output to deal with
</aside>

Let's just test this out quickly:

```{r bbtex_01_02, eval=TRUE, echo=TRUE, cache=TRUE}
lec_no <- 1
get_lec_bibtex(lec_num = lec_no)
```

Great - looks like it is working as required with the correct string padding
in the lecture number in the pdf filename!

<aside>
**Note:** We used the `stringr` `str_pad` to convert `1` to `"01"`
</aside>

### Apply to all lectures using `purrr`

Let's finish this by creating all the entries using `purrr`:

```{r bbtex_01_03, eval=TRUE, echo=TRUE, cache=TRUE}
lec_nums <- c(1, 30)
lec_nums %>% 
  purrr::map_chr(.x = ., .f = ~get_lec_bibtex(lec_num = .x)) %>% 
  cat(., sep = "\n\n")
```

Yay - this works as expected! We can now paste into BibTeX 
as required.

Note that we only created it for lectures 1 and 30 for easy 
scrolling. But for all lectures we can just 
replace `c(1, 30)` with `1:30` in the above code.

## Conclusion

This post was for me to document and serve as a guide to automating 
a couple of fun text-based tasks that I came across in my work
(and social life!). Using the `tidy` framework can be a fun way
to solve these tasks (but certainly not the only way in `R`). 
Have fun playing around with the above and post in the comments
any feedback you have `r emo::ji("thumbsup")`.

Stay tuned for more blogposts solving more such tasks.

## Acknowledgments {.appendix}

I'd like to thank my brother Salil Shrotriya for helping
me create the preview image for this post using his cool
photoshop skills. The hex sticker `png` files were sourced
from [here](https://github.com/rstudio/hex-stickers/raw/master/SVG/)