---
title: "Tidyverse Fun - Part 1"
description: |
  First part in a series of doing useful tasks with
  the `Tidyverse`
categories:
  - tidyverse
  - rstats
author:
  - name: Shamindra Shrotriya
    url: https://www.shamindras.com/
date: 07-15-2019
base_url: https://www.shamindras.com/
preview: images/logo-01.jpg
slug: shrotriya2019tidyfunpt1
bibliography: ../../refs.bib
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

```{r setup_01_01, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1: Generating Oxford Comma Triples

### The central problem

Based on a fun conversation with my statistics cohort over dinner we got to
discussing the famous *Oxford Comma* [@wiki_serial_comma] (or *Serial Comma*
depending on your persuasion). I've never really adopted the use but my friends
made a compelling argument on it's apparent general lack of ambiguity when
applied appropriately.

We will use the Oxford comma on the famously *ambiguous* phrase (here used
without the Oxford Comma before _leaves_):

> Eats, shoots and leaves

After adding in the Oxford Comma this would become:

> Eats, shoots, and leaves

**Goal:** A fun experiment would be to generate *all permutations* of this
phrase with and without the Oxford Comma using `R` and specifically the
`tidyverse` packages.

### Generating all word-triple permutations the `tidy` way

Let's load our required packages

```{r setup_oxc_01_01, echo=TRUE, cache=TRUE}
library(magrittr)
library(tidyverse)
library(glue)
```

Let's also define our unique global word values used to construct the required
phrases:

```{r setup_oxc_01_02, echo=TRUE, cache=TRUE}
WORD_VALS <- c("eats", "shoots", "leaves")
```

Generate all unique 3-word permutations _without replacement_ from the three
unique words. We'll create a helper function to check that a vector of words is
unique.

```{r setup_oxc_01_03, echo=TRUE, cache=TRUE}
is_unq_perm <- function(word1, word2, word3){
    words_vec <- c(word1, word2, word3)
    base::return(length(words_vec) - length(unique(words_vec)) == 0)
}
```

We can now simply generate every possible triple _with replacement_ using the
`tidyr::crossing` function.  We proceed to _filter_ these $3^3 = 27$ triples
for unique triples using our `is_unq_perm` helper function applied _row-by-row_
using `purrr::pmap_lgl`. The `_lgl` simply returns a `TRUE/FALSE` logical value
as intended by the applied function.

<aside> __Note:__ The `tidyr::crossing` generates a *Cartesian product* of all
the 3 word triples, very handy</aside>

```{r setup_oxc_01_04, echo=TRUE, cache=TRUE}
# Generate the unique word-triples
all_perms <- tidyr::crossing(word1 = WORD_VALS,
                             word2 = WORD_VALS,
                             word3 = WORD_VALS) %>%
                dplyr::mutate(.data = .,
                              is_unq_perm = purrr::pmap_lgl(.l = .,
                                                            is_unq_perm)) %>%
                dplyr::filter(.data = ., is_unq_perm) %>%
                dplyr::select(-is_unq_perm)

# Display output in a nice centered table
all_perms %>%
  knitr::kable(x = ., align = 'c',
               col.names = c("Word 1",
                             "Word 2",
                             "Word 3"))
```

Great - that part is done! Now we just need to generate for each triple of
words an oxford comma and non-oxford comma version. This is done easily using
the amazing `glue` package as seen below:

```{r setup_oxc_01_05, cache=TRUE, echo=TRUE}
exprs <- all_perms %>%
          dplyr::mutate(non_oxford_comma =
                          glue::glue_data(.x = .,
                          "{word1}, {word2} and {word3}"),
                        oxford_comma =
                          glue::glue_data(.x = .,
                          "{word1}, {word2}, and {word3}")) %>%
          dplyr::select(non_oxford_comma, oxford_comma)
```

We can display the side-by-side output of the Non-Oxford Comma vs. Oxford comma
for the $6$ generated triples as follows:

```{r setup_oxc_01_06, cache=TRUE, echo=TRUE}
# Display output in a nice centered table
exprs %>%
  knitr::kable(x = .,
               align = 'c',
               col.names = c("Non-Oxford Comma",
                             "Oxford Comma"))
```

So there you have it. Have fun generating your own version of Oxford Comma
triples to engage in civil discussions with your fellow grammar focused friends
`r emo::ji("smile")`.

## Task 2: Generating numbered BibTeX entries

### The central problem

In this case I needed to generate several BibTeX entries of the form:

```markup
@misc{doe2019_lec1,
author        = {Doe, John},
title         = {Lecture Note 1 - STAT10A},
month         = {March},
year          = {2018},
url           = {https://statschool/~doe/stats10A/Lectures/Lecture01.pdf},
}
```

As it can be seen the lectures are numbered sequentially and change in the main
BibTeX `id`, the `title`, and the `url` field.

Specifically I needed to construct 30 such sequential entries for lectures
`1-30`. Rather than do this manually, I realized that this would be fun
scripting exercise with using the `tidyverse` packages `glue`, `purrr`, and
`stringr`.

**Goal:** Create 30 such BibTeX entries and print to the console to
directly-copy paste to my BibTeX file.

### The `tidy` approach

First step is to write a function that takes a lecture number (integer) as an
input and then outputs a single BibTeX entry for that lecture.

```{r bbtex_01_01, eval=TRUE, echo=TRUE, cache=TRUE}
# Generate BibTeX entry for a single lecture number
get_lec_bibtex <- function(lec_num){
  # Get the 2 character padded lecture number i.e. 1 -> "01"
  lec_num_pad <- stringr::str_pad(string = lec_num, width = 2,
                                  side = "left", pad = "0")

  # Construct the BibTeX entry
  out_bbtex_str <- glue::glue(
    "@misc{doe2019_lec<lec_num>,
    author = {Doe, John},
    title  = {Lecture Note <lec_num> - STAT10A},
    month  = {March},
    year   = {2018},
    url    = {https://www.hpg/~doe/st10A/lecs/lec<lec_num_pad>.pdf}}",
    .open = "<",
    .close = ">")

  base::return(out_bbtex_str)
}
```

Note that by default `glue` allows you to substitute input text in between `{`
and `}` markers. However BibTeX entries *already have* literal default `{}`
tags that we need to include in our function output. Rather than escaping them
the `glue` package conveniently allows us to change the default opening and
closing markers `r emo::ji("hundred")`! We simply set these to be angle
brackets `< >` using the `.open` and `.close` options above.

<aside> __Note:__ Luckily we don't have _literal_ angle brackets in our BibTeX
output to deal with here</aside>

Let's just test this out quickly:

```{r bbtex_01_02, eval=TRUE, echo=TRUE, cache=TRUE}
lec_no <- 1
get_lec_bibtex(lec_num = lec_no)
```

Great - looks like it is working as required with the correct string padding in
the lecture number in the pdf filename!

<aside> **Note:** We used the `stringr` `str_pad` to convert `1` to `"01"`
</aside>

### Apply to all lectures using `purrr`

Let's finish this by creating all the entries using `purrr`:

```{r bbtex_01_03, eval=TRUE, echo=TRUE, cache=TRUE}
lec_nums <- c(1, 30)
lec_nums %>%
  purrr::map_chr(.x = ., .f = ~get_lec_bibtex(lec_num = .x)) %>%
  cat(., sep = "\n\n")
```

Yay - this works as expected! We can now paste into BibTeX as required.

Note that we only created it for lectures 1 and 30 for easy scrolling. But for
all lectures we can just replace `c(1, 30)` with `1:30` in the above code.

## Conclusion

This post was for me to document and serve as a guide to automating a couple of
fun text-based tasks that I came across in my work (and social life!). Using
the `tidy` framework can be a fun way to solve these tasks (but certainly not
the only way in `R`). Have fun playing around with the above and please post in
the comments any questions/feedback you may have `r emo::ji("thumbsup")`.

Stay tuned for more blogposts solving more such tasks.


```{r, child="../../data/rmds/acknowledgment_preview_tidyfun_salil.Rmd"}
```
