---
title: "About Me"
description: |
  A brief bio and welcome to my site
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<center>
```{r main-src-poster, echo=FALSE}
IMAGES_PATH <- here::here("data", "images")
knitr::include_graphics(path = fs::path_join(parts = c(IMAGES_PATH,
                                                       "about-profile-circular-scaled-75.png")))
```
</center>

Hi there! I’m **Shamindra Shrotriya**, a graduate student in the [Department of
Statistics and Data Science](http://www.stat.cmu.edu/){target="_blank"} at the
wonderful [Carnegie Mellon University](https://www.cmu.edu/){target="_blank"}.
I’ve decided to start blogging to document my learning and development in
statistics as a graduate student.

I’d like it to be a fun place to document interesting things I like to read
about in the statistics and machine learning space (**statistical
theory/methodology**, **research**, **rstats**, **python** ...) as well as
anything else I am generally into e.g. **books**, **sports** etc.

Feel free to pull up a chair, leave a comment, and join me so that we can
explore together.

My CV can be found [here](data/pdfs/CV.pdf){target="_blank"}.

```{r, child="./data/rmds/favicons_shamindra_shrotriya.Rmd"}
```

## Acknowledgments {.appendix}

I'd like to thank Salil Shrotriya for taking the profile pic.
